Is structure based drug design ready for selectivity optimization?

### 2018-10-30 manuscript.tex
checking things in before the train

Since the FDA approval of imatinib in 2001, targeted small molecule kinase inhibitors (SMKIs) have become a major class of therapeutics in treating cancer and other diseases. Currently, there are 38\citep{fda-approved-kinase-inhibitors} approved SMKIs, and it is estimated that kinase targeted therapies account for as much as 50\% of current drug development~\citep{Santos:Nat.Rev.DrugDiscov.:2016}, with many more compounds currently in clinical trials. While there have been a number of successes, design of selective kinase inhibitors remains a challenge. Selectivity is particularly difficult for kinases: there are close to 518 protein kinases~\citep{Volkamer2015-jx,Manning2002-cw} with a highly conserved ATP binding site that is targeted by the majority of small molecule kinase inhibitors~\citep{Wu2015-oq}. While these inhibitors do adopt several different binding modes~\citep{Cowan-Jacob2007-rn,Seeliger2007-jn,Huse2002-ml,Harrison2003-ct}, previous work has shown that both Type I (binding to the active, DFG-in conformation) and Type II (binding to the inactive, DFG-out conformation) inhibitors display a wide range of selectivities~\citep{Anastassiadis2011-sm,Davis:Nat.Biotechnol.:2011}, often binding a number of other targets unselectively in addition to a primary target. Even inhibitors approved by the FDA, often the result of extensive drug development programs, can bind to a large number of off-target kinases~\citep{Klaeger2017-jr}.

@@

\begin{figure}
\begin{fullwidth}
\begin{centering}
\includegraphics[width=1.0\linewidth]{figures/figure4.png}
% Need at least one blank line for centering to work.
\end{centering}
\caption{
\label{fig:figure-4}
{\bf Relative free energy calculations can accurately predict potency, but show larger errors for selectivity predictions.} \\
({\bf A}) Single target potencies for CDK2, ERK2 and CDK9 from both the Shao and Blake datasets, with experimental values on the X-axis and calculated values on the Y-axis. Each black point corresponds to a ligand for a given target. All values are shown in units of kcal/mol. The horizontal error bars show the assumed experimental uncertainty of 0.3 kcal/mol\citep{BROWN2009420}. To better highlight outliers that are unlikely due simply to forcefield errors, we presume the forcefield error ($\sigma_\mathrm{FF} \approx$ 0.9 kcal mol$^{-1}$~\cite{Harder:J.Chem.TheoryComput.:2016}) also behaves as a random error. We show the total estimated statistical and forcefield error ($\sqrt{\sigma_\mathrm{FF}^2 + \sigma_\mathrm{calc}^2}$) as vertical error bars. The black line indicates agreement between calculation and experiment, while the gray shaded region represent 1.36 kcal/mol (or 1 log unit) error. The MUE and RMSE are shown on each plot with bootstrapped 95$\%$ confidence intervals.
({\bf B}) The indicated $\Delta\Delta$G is shown for each ligand. 
}
\end{fullwidth}
\end{figure}

>>>>

With the FDA approval of imatinib in 2001, targeted small molecule kinase inhibitors (SMKIs) have become a major class of therapeutics in treating cancer and other diseases. Currently, there are 43\citep{fda-approved-kinase-inhibitors} approved SMKIs, and it is estimated that kinase targeted therapies account for as much as 50\% of current drug development~\citep{Santos:Nat.Rev.DrugDiscov.:2016}, with many more compounds currently in clinical trials. While there have been a number of successes, design of selective kinase inhibitors remains a challenge. Selectivity is particularly difficult for kinases: there are more than 518 protein kinases~\citep{Volkamer2015-jx,Manning2002-cw} with a highly conserved ATP binding site that is targeted by the majority of small molecule kinase inhibitors~\citep{Wu2015-oq}. While these inhibitors do adopt several different binding modes~\citep{Cowan-Jacob2007-rn,Seeliger2007-jn,Huse2002-ml,Harrison2003-ct}, previous work has shown that both Type I (binding to the active, DFG-in conformation) and Type II (binding to the inactive, DFG-out conformation) inhibitors display a wide range of selectivities~\citep{Anastassiadis2011-sm,Davis:Nat.Biotechnol.:2011}, often binding a number of other targets unselectively in addition to a primary target. Even inhibitors approved by the FDA, often the result of extensive drug development programs, can bind to a large number of off-target kinases~\citep{Klaeger2017-jr}.

@@

\begin{figure}
\begin{fullwidth}
\begin{centering}
\includegraphics[width=1.0\linewidth]{figures/figure4.png}
% Need at least one blank line for centering to work.
\end{centering}
\caption{
\label{fig:figure-4}
{\bf Relative free energy calculations can accurately predict potency, but show larger errors for selectivity predictions.} \\
Single target potencies and selectivities for CDK2/ERK2 from the Blake datasets (\emph{top}), and CDK2/CDK9 (\emph{bottom}) from the Shao datasets. The experimental values are shown on the X-axis and calculated values on the Y-axis. Each data point corresponds to a ligand for a given target. All values are shown in units of kcal/mol. The horizontal error bars show the assumed experimental uncertainty of 0.3 kcal/mol\citep{BROWN2009420}. To better highlight outliers that are unlikely due simply to forcefield errors, we presume the forcefield error ($\sigma_\mathrm{FF} \approx$ 0.9 kcal mol$^{-1}$~\cite{Harder:J.Chem.TheoryComput.:2016}) also behaves as a random error. We show the total estimated statistical and forcefield error ($\sqrt{\sigma_\mathrm{FF}^2 + \sigma_\mathrm{calc}^2}$) as vertical blue error bars. The black vertical error bars correspond to the statistical error ($\sigma_{calc}$). The black line indicates agreement between calculation and experiment, while the gray shaded region represent 1.36 kcal/mol (or 1 log unit) error. The MUE and RMSE are shown on each plot with bootstrapped 95$\%$ confidence intervals.
}
\end{fullwidth}
\end{figure}

\begin{figure}
\begin{fullwidth}
\begin{centering}
\includegraphics[width=1.0\linewidth]{figures/figure5.png}
% Need at least one blank line for centering to work.
\end{centering}
\caption{
\label{fig:figure-5}
{\bf Relative free energy calculations can accurately predict potency, but show larger errors for selectivity predictions.} \\
}
\end{fullwidth}
\end{figure}
